Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2020-12-30 papers/thesis-kirstie/thesis.tex
edits

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a repulsive 
potential. This model, called the hard-sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
fluid which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials to create the Barker-Henderson model 
fluid~\cite{barker1967perturbation}. In this model, a soft-sphere fluid is 
modeled by using a hard-sphere fluid with hard spheres of larger radii 
at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid.

ADD MORE HERE TO COME UP TO DATE!!
\newline

While models for particular fluids were being created, powerful methods 
to describe the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
%\textcolor{red}{
In 1958, Percus-Yevick introduced an alternative method 
for the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
and the corresponding equilibrium number density profile for 
inhomogeneous fluids. To use the classical Density Functional Theory (cDFT) 
approach, an expression for the free energy 
as a functional of an average number density that varies spatially 
must be constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
the lowest term in the virial expansion 
of the excess chemical potential for a hard-sphere fluid can 
be expressed in terms of a sum of weighted densities using a weight 
function of the form of a backward Heavyside step function (shown below) 
that can be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere.
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick.
Tarazona later improved on the 
method by introducing a tensor version of one of the energy density terms 
in the functional~\cite{tarazonaphi3, 
santos2012phi3}, and in 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).

Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
In this paper, classical Density Functional 
Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to fit the 
functional to the soft-sphere WCA fluid.
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

Freezing is governed by entropy, which the universe %, or any closed system, 
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), an abrupt phase transition occurs, that is, the liquid 
freezes. The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

@@

%In order to apply classical Density Functional Theory to a system that has
%a fixed temperature, volume, and number of particles, a functional for 
%the Helmholtz free energy must first be formed. An expression for the ideal 
%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
%but forming an expression for the excess Helmholtz free
%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
%being modeled can be challenging. 
%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
%expression for the excess Helmholtz free energy of a soft-sphere fluid
%that does not depend on explicitly taking into account the interaction 
%that each sphere with every other sphere in the fluid (as required by 
%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
%originally developed by 
%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
%for the hard-sphere fluid, and which expanded 
%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
%to inhomogeneous, or non-uniform hard-sphere fluids.

@@

Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
%and provides an expression for the excess Helmholtz free energy of an 
%inhomogeneous 
%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
%
%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
%is an extension of FMT and provides an expression for the excess Helmholtz
%free energy of fluids with soft potentials like that of the WCA fluid. 
%It will be helpful to continue looking at FMT before addressing SFMT.
%FMT begins 
by noticing that a quantity $G_{3}$ can be defined such that
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
where $G_{3}$ is the volume of a sphere. 
Likewise, by defining quantities
\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=R}\end{equation}
\begin{equation}{G_{0}=1}\end{equation}
where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
geometric measures of a sphere, and are what give rise to the name
Fundamental Measure Theory. Introducing a set of variables
%weighted densities $n_{i}=nG_{i}$ 
given by
\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
Equation~\ref{FexfromSPTinthermsofG} can be written as
\begin{align}\label{FexfromSPT}
    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
\end{align}
which has the form 
\begin{align}
    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
\end{align}
%where 
\begin{align}
   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
\begin{align}
    n_{i}=nG_{i}
\end{align}
These weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extension to a system with a spatially varying number 
density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
excess Helmholtz free energy is written as
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
Using a backwards Heavyside step function defined as
\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
the following expressions are obtained
\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
where the weighted densities can now be expressed as
\begin{align}
    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
\end{align} 
or for a spatially varying number density $n(\vec{r})$,
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
The weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - a "liquid" region at lower densities, and a solid 
region at higher densities. 
A phase transition between a liquid and a solid is accompanied by an 
abrupt change in the number density, and so there is a gap between the 
purely liquid and the purely solid regions. Within the gap, or coexistence
region, is found mixtures of various amounts of liquid and solid each of
which have number densities that are fixed at the values they had at the 
phase transition.

To construct the phase diagram, the liquid and solid phase transition number densities
are found

%Maxwell's double tangent construction can be used to find 
%these number densities given plots of

from the liquid and the solid Helmholtz 
free energies and inverse densities at a given temperature, as
discussed in Section 2.2.

@@

\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

Calculate $\widetilde{w}_0(k)$ 
\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}

>>>>

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a repulsive 
potential. This model, called the hard-sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms.

In more recent times, other model fluids have been developed.
Zwanzig introduced the square-well 
fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
In this model, 
%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
hard spheres are given larger radii 
at low temperatures, and smaller radii at high temperatures. 
In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid. 
Models continue to be developed to characterize fluids with particular 
properties, or to address particular problems like surface tension, 
interfaces, adsorption, wetting or confinment.
\textcolor{red}{A review of model fluids, and problems studied more 
recently is given in ...}

While models for particular fluids were being created, powerful methods 
to predict the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
of the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
%, and the corresponding equilibrium number density profile 
of a classical inhomogeneous fluid. 
Density functional theory was further developed for inhomogeneous classical fluids
by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
To use classical Density Functional Theory (cDFT), an expression for the free energy 
as a functional of a spatially varying average single-particle number density 
%that varies spatially 
is constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
%the lowest term in the virial expansion of the excess chemical potential 
the lowest order term in the virial expansion of the excess chemical 
potential of a hard-sphere fluid can 
be expressed in terms of a weighted density using a weight function 
%using a weight 
%function 
of the form of 
a backward Heaviside step function}
%\begin{equation}
   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
%\end{equation}
that can itself be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere. 
For a hard-sphere fluid with spheres of two different sized radii, 
$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick. 
However, it did not predict freezing due to one term in the expression that
diverged when the fluid was highly confined.  
Tarazona later solved this problem by introducing a tensor version of 
one of the energy density terms 
in the functional~\cite{tarazonaphi3, santos2012phi3} 
which took into account the non-spherical, and thus 
non-symmetric, shape of the cavity in which the fluid is confined 
while also yielding the correct excess free energy for the fluid as its 
confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
to a two-dimensional plane (as in a thin-film), to a one-dimensional 
line (as in a fluid confined to a capillary tube), to a zero-dimensional 
dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.

In 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).
Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
\textcolor{red}{For more recent developments in classical 
Density Functional Theory see (2020), Evan's review~\cite{evansreview} (2016), 
and Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013).}

FREEZING in literature :
\textcolor{red}{Classical density functional theory was used
to ... (paper from 2010), but not until recently has there been good success 
(other 23 papers...), and in all these the hard-sphere fluid has been 
used as the reference system.} 
In this paper, classical Density Functional 
Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to fit the 
functional to the soft-sphere WCA fluid.
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

Freezing of simple fluids is governed by entropy, which the universe
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), a phase transition occurs, that is, the liquid 
freezes. 
Both the entropy and the volume (and thus the density) change abruptly 
at the phase transition.
The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

@@

\textcolor{blue}{
Rosenfeld begins the development of FMT by considering the lowest order 
term in the virial expansion of the excess chemical potential.
Taking the partial derivative of the free energy in the low density limit given 
by Equation~\ref{eq:Fexcess} gives
\begin{align}
   \mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
            &= k_BT\int -f(r)~n~d\vec r
\end{align}
For a hard-sphere fluid (simplified in this account by considering only 
equal sized spheres), the Mayer function, $f(r)$ can be written in terms 
of a backward Heaviside step function
\begin{align} 
    f(r) = -\theta(|\vec r|-2R)
\end{align}
where and $\vec r$ is the distance between the centers of two spheres.
The low-density excess free energy given by Equation~\ref{eq:Fexcess} 
can then be written as a weighted density.
\begin{align}
     F_{excess}=k_BT\left(\int n^2(\vec r)\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
\end{align} 
where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
Rosenfeld recognized that the Heaviside function describes mathmetically 
the exclusion volume of a pair of spheres, one of the two variables 
(the other one being the "overlap volume") in which the total 
excess chemical potential can be expanded. He also noted that the 
Heaviside function can uniquely be deconvolved into a set of weight functions
related to the geometry of a sphere.
\begin{align}
    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
\end{align} 
where again this expression is simplified in this account by considering only 
equal sized spheres. 
The excess free energy in the low density limit then takes the form
\begin{align}
     F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
\end{align}
%which can now be expressed in terms of a set of weighted densities
\begin{align}
     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
\end{align}
}
%In order to apply classical Density Functional Theory to a system that has
%a fixed temperature, volume, and number of particles, a functional for 
%the Helmholtz free energy must first be formed. An expression for the ideal 
%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
%but forming an expression for the excess Helmholtz free
%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
%being modeled can be challenging. 
%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
%expression for the excess Helmholtz free energy of a soft-sphere fluid
%that does not depend on explicitly taking into account the interaction 
%that each sphere with every other sphere in the fluid (as required by 
%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
%originally developed by 
%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
%for the hard-sphere fluid, and which expanded 
%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
%to inhomogeneous, or non-uniform hard-sphere fluids.

@@

Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
%and provides an expression for the excess Helmholtz free energy of an 
%inhomogeneous 
%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
%
%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
%is an extension of FMT and provides an expression for the excess Helmholtz
%free energy of fluids with soft potentials like that of the WCA fluid. 
%It will be helpful to continue looking at FMT before addressing SFMT.
%FMT begins 
by noticing that a quantity $G_{3}$ can be defined such that
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
where $G_{3}$ is the volume of a sphere. 
Likewise, by defining quantities
\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=R}\end{equation}
\begin{equation}{G_{0}=1}\end{equation}
where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
geometric measures of a sphere, and are what give rise to the name
Fundamental Measure Theory. Introducing a set of variables
%weighted densities $n_{i}=nG_{i}$ 
given by
\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
Equation~\ref{FexfromSPTinthermsofG} can be written as
\begin{align}\label{FexfromSPT}
    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
\end{align}
which has the form 
\begin{align}
    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
\end{align}
%where 
\begin{align}
   \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
\begin{align}
    n_{i}=nG_{i}
\end{align}
These weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extension to a system with a spatially varying number 
density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
excess Helmholtz free energy is written as
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
Using a backwards Heaviside step function defined as
\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
the following expressions are obtained
\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
where the weighted densities can now be expressed as
\begin{align}
    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
\end{align} 
or for a spatially varying number density $n(\vec{r})$,
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
The weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - a "liquid" region at lower densities, and a solid 
region at higher densities. 
A phase transition between a liquid and a solid is accompanied by an 
abrupt change in the number density, and so there is a gap between the 
purely liquid and the purely solid regions. Within the gap, or coexistence
region, is found mixtures of various amounts of liquid and solid each of
which have number densities that are fixed at the values they had at the 
phase transition. 
To construct the phase diagram, the liquid and solid phase transition number densities
are found 
%Maxwell's double tangent construction can be used to find 
%these number densities given plots of 
from the liquid and the solid Helmholtz 
free energies and inverse densities at a given temperature, as
discussed in Section 2.2.

@@

\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

Calculate $\widetilde{w}_0(k)$ 
\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}

\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
